Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-06 ms.tex
randome2

As the total heating rate declines with time, after some point the
flow becomes thermally unstable according to the criterion
$(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}} \lesssim 10$
(eq.~[\ref{eq:cooling3}]), as shown by dashed lines in
Fig.~\ref{fig:vwSourcesImp}.  Compton heating is neglected in
calculating thermal stability because unlike local stellar feedback
mechanisms, SMBH feedback may not be capable of truly stabilizing the
flow given its inability to respond instantaneously to local changes
in the properties of the gas.  Thermally unstable flow is present
throughout the single starburst case except at very early times,
$\tau_{\star} \lesssim 10^7$ yr, and at late times $\tau_{\star}
\gtrsim 10^{8}$ yr in the high $M_{\bullet}$ case.

@@

Figure~\ref{fig:vwSources} shows $\vwO(M_{\bullet})$ from each heating
source (stars, MSPs, SNe Ia, and black hole feedback) calculated for
the average star formation history of galaxies containing a given
black hole mass.  The younger stellar populations characterizing galaxies
with $\Mbh\lsim 3\times 10^{8} \Msun$ are dominated by stellar wind, with $v_{w}^{\star} \gtrsim 700$ km s$^{-1}$.  Only for the most massive galaxies, with $\Mbh\gsim 3\times 10^{8} \Msun$, does the lack
of young stellar populations significantly reduce the role of stellar
wind heating.  But for these massive galaxies, Ia heating contributes a comparable level of heating $v_{w}^{\rm Ia} \gtrsim 600$ km s$^{-1}$.  However, for low mass galaxies Ia heating does not contribute to the total heating rate because the Ia radius exceeds the nominal stagnation radius (bottom panel).  Finall, MSP and Compton heating is found to be negligible for all black hole masses.  Strikingly, {\it we find that galactic nuclei experiencing the same
  average star formation history as their host galaxies produce thermally stable flows across almost all black hole masses}.  
We must caveat, however, that for high $M_{\bullet}$ the total effective wind velocity is only a factor of 2-3 times higher than the stellar velocity dispersion ($\zeta \sim \zeta_c$), such that thermal stability is marginal, especially for core galaxies ($\S\ref{sec:mdot}$).

Realistic variations (``burstiness") in the star formation history
will also often result in lower heating rates (Appendix
\ref{app:windheat}).  Figure \ref{fig:NickPlot2} shows the heating
rate as a function of average black hole mass for non-fiducial cases
in which the current ($z = 0$) star formation is suppressed by a
factor $\iota$ from its average $z = 0$ value, for a characteristic
timescale $\delta t_{\star}$.  For $\delta t_{\star} \lesssim 10^{7}$
yr, we see that $v_{\rm w}^{\star}$ is reduced by at most a factor of
$\approx 2$ from its average value, even for huge drops in the star
formation rate ($\iota \sim 10^{-3}$).  However, burstiness in the
star formation history over longer timescales can suppress heating
more significantly.  When $\delta t_{\star} \gtrsim 10^8~{\rm yr}$,
$v_{\rm w}^\star$ becomes very sensitive to $\iota$: at $\iota=0.1$,
$v_{\rm w}^\star \approx 500~{\rm km~s}^{-1}$ in most galaxies, but
for smaller $\iota$, $v_{\rm w}^\star \approx 200~{\rm km~s}^{-1}$ (an
exception to this is in galactic nuclei with $M_\bullet \gtrsim
5\times 10^8~M_\odot$, where heating is stabilized by SN Ia).  Our use
of average cosmic star formation histories is appropriate provided
local variations are either on short timescales ($\delta t_{\star}
\lesssim 10^7~{\rm yr}$), or limited in magnitude ($\iota \gtrsim
0.1$).  If both of these conditions are violated, then the effective
heating rates for all but the most massive galaxies (where Ia
explosions dominate) will fall by a factor $\approx 5$ from our
fiducial volumetric averages, taken from \citet{MosterNaab+:2013a}.

>>>>

As the total heating rate declines with time, after some point the
flow becomes thermally unstable according to the criterion
$(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}} \lesssim 10$
(eq.~[\ref{eq:cooling3}]), as shown by dashed lines in
Fig.~\ref{fig:vwSourcesImp}.  Compton heating is neglected in
calculating thermal stability because, unlike local stellar feedback
mechanisms, SMBH feedback may not be capable of truly stabilizing the
flow given its inability to respond instantaneously to local changes
in the properties of the gas.  Thermally unstable flow is present
throughout the single starburst case except at very early times,
$\tau_{\star} \lesssim 10^7$ yr, and at late times $\tau_{\star}
\gtrsim 10^{8}$ yr in the high $M_{\bullet}$ case.

@@

Figure~\ref{fig:vwSources} shows $\vwO(M_{\bullet})$ from each heating
source (stars, MSPs, SNe Ia, and black hole feedback) calculated for
the average star formation history of galaxies containing a given
black hole mass.  The younger stellar populations characterizing galaxies
with $\Mbh\lsim 3\times 10^{8} \Msun$ are dominated by stellar winds, with $v_{w}^{\star} \gtrsim 700$ km s$^{-1}$.  Only for $\Mbh\gsim 3\times 10^{8} \Msun$ does the lack
of young stellar populations significantly reduce the role of stellar
wind heating.  But for these massive galaxies, that Ia radius is sufficiently small for Ia heating to contribute a comparable level of heating, $v_{w}^{\rm Ia} \gtrsim 600$ km s$^{-1}$ (for low mass galaxies Ia heating does not contribute to the total heating rate because $r_{\rm Ia} > \rs$; bottom panel).  Strikingly, we find that galactic nuclei experiencing the same
  average star formation history as their host galaxies produce thermally stable flows across almost all black hole masses.

An important caveat, however, is that $\vw$ is only a few times higher than the stellar velocity dispersion, i.e. $\zeta \sim \zeta_c$.  This implies that the true stagnation radius could be larger than our analytic estimates, potentially resulting in thermal instability, especially for core galaxies with large break radii ($\S\ref{sec:mdot}$).

Realistic variations (``burstiness") in the star formation history
will also often produce lower stellar wind heating rates (Appendix
\ref{app:windheat}).  Figure \ref{fig:NickPlot2} shows the heating
rate as a function of average black hole mass for non-fiducial cases
in which the current ($z = 0$) star formation is suppressed by a
factor $\iota$ from its average $z = 0$ value, for a characteristic
timescale $\delta t_{\star}$.  For $\delta t_{\star} \lesssim 10^{7}$
yr, we see that $v_{\rm w}^{\star}$ is reduced by at most a factor of
$\approx 2$ from its average value, even for huge drops in the star
formation rate ($\iota \sim 10^{-3}$).  However, burstiness in the
star formation history over longer timescales can suppress heating
more significantly.  When $\delta t_{\star} \gtrsim 10^8~{\rm yr}$,
$v_{\rm w}^\star$ becomes very sensitive to $\iota$: at $\iota=0.1$,
$v_{\rm w}^\star \approx 500~{\rm km~s}^{-1}$ in most galaxies, but
for smaller $\iota$, $v_{\rm w}^\star \approx 200~{\rm km~s}^{-1}$ (an
exception to this is in galactic nuclei with $M_\bullet \gtrsim
5\times 10^8~M_\odot$, where heating is stabilized by SN Ia).  Our use
of average cosmic star formation histories is appropriate provided
local variations are either on short timescales ($\delta t_{\star}
\lesssim 10^7~{\rm yr}$), or limited in magnitude ($\iota \gtrsim
0.1$).  If both of these conditions are violated, then the effective
heating rates for all but the most massive galaxies (where Ia
explosions dominate) will fall by a factor $\approx 5$ from our
fiducial volumetric averages, taken from \citet{MosterNaab+:2013a}.
